### 2016-09-15 paper/steps.tex

\begin{tikzpicture}[xscale=0.7]

\tikzstyle{state} = [rectangle, draw=black,fill=white,rounded corners=1mm, inner sep=2mm,minimum width=14pt,font=\scriptsize,text=black]

\tikzstyle{sdfgnode} = [circle,draw=black,fill=white,inner sep=0,minimum width=14pt,font=\scriptsize,text=black]%
\tikzstyle{sdfgdata} = [sdfgnode, inner sep=1pt,fill=white]%
\tikzstyle{sdfgmem} = [sdfgnode, rectangle, rounded corners=4pt, inner sep=2pt,fill=white]%
\tikzstyle{sdfgarc} = [->,minimum width=2pt,>=latex,sloped]%

\node[state] (inplang) at (0,0) {Specification};
\node[state] (csdfgraph) at (5,0) {CSDF graph};
\node[state] (hardware) at (10,1) {Hardware};
\node[state] (software) at (10,-1) {Software};

\draw[sdfgarc] (inplang) -- node[above, font=\scriptsize] {Parsing} (csdfgraph);
\draw[sdfgarc] (csdfgraph) -- node[above, font=\scriptsize] {Code gen.} (hardware);
\draw[sdfgarc] (csdfgraph) -- node[above, font=\scriptsize] {Code gen.} (software);

\draw[sdfgarc] (csdfgraph)
    .. controls ($(csdfgraph) + (-50:1)$) and ($(csdfgraph) + (0.5,-1)$) .. ($(csdfgraph) + (0, -1)$)
    node[below, font=\scriptsize] {Transformations}
    .. controls ($(csdfgraph) + (-0.5,-1)$) and ($(csdfgraph) + (-130:1)$) ..
    (csdfgraph);

\end{tikzpicture}